Rising from the Ashes: A Metallicity-Dependent Star Formation Gap Splits the Milky Way's $\alpha$-Sequences

### 2025-02-13 paper/ms.tex
Update conc and ack

In this work, using idealized merger simulations, we have shown that (1) a starburst, followed by (2) a brief ($\sim300\,\Myr$) quiescent period, followed by (3) rejuvenation can lead to a bimodal abundance structure. We found that the starburst-to-quiescent period allowed for prodigious Fe-enrichment without any contemperaneous $\alpha$-enrichment. This sequence of events, somewhat resembling the life cycle of a phoenix, would predict that for stars at a given \FeH{}, there would be a $\sim300\,\Myr$ gap in stellar ages at $\sim8\,\Gyr$. Furthermore, stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly later ($\sim1\,\Gyr$).

The bimodality in the \alphaFe{}-\FeH{} plane of stellar abundances points to deeper, unobserved structure. That structure is a record of the complicated processes that gave rise to the Galaxy. In the absence of highly precise stellar ages, that history must be largely inferred. There are many ways in which a galaxy could end up with the coarse-grained structure we can observe in the present day. We have argued in this work that a brief, starburst-driven quiescent phase is one of them.

\begin{acknowledgements}
We would like to thank Megan Bedell, Christopher Carr, Vedant Chandra, Charlie Conroy, Drummond B. Fielding, Lars Hernquist, Federico Marinacci, Rohan P. Naidu, Melissa K. Ness, Minjung Park, Vadim A. Semenov, and Turner Woody for helpful discussions. We would like to thank Filippo Barbani for kindly providing his version of \texttt{MakeNewDisk}. A.B. would like to thank Todd Phillips for helpful discussions.

>>>>

\edit1{In this work, we have shown that a brief ($\sim300\,\Myr$) period of halted star formation in a narrow \FeH{} bin is capable of producing a bimodal distribution in \alphaFe{} at that \FeH{}. This proposal requires that the \alphaFe{} of gas within the galaxy is decreasing with time so that the gap in star formation translates to a gap in \alphaFe{}. A global quiescent period can satisfy these constraints, but is not necessary. We demonstrate the plausibility of this scenario using a grid of idealized merger simulations with slightly varied orbital parameters -- however, it is also not necessary that a merger triggers the \FeH{}-dependent halt in star formation.}

\edit1{This scenario leads to the natural prediction that for stars occupying a narrow bin in \FeH{} where the bimodality is present ($\FeH\lesssim-0.2$), there should be a gap in ages for $\sim8\,\Gyr$ old stars with a width of $\sim300\,\Myr$. Currently, the best age estimates for such stars have errors of $\sim1\,\Gyr$. However, future observations targeting \textit{relative} ages of such stars might be able to achieve the necessary precision. Our proposed mechanism may operate in many external galaxies, whether or not these \FeH{}-dependent metallicity gaps are merger-induced.}

\begin{acknowledgements}
We would like to thank Megan Bedell, Christopher Carr, Vedant Chandra, Charlie Conroy, Drummond B. Fielding, Lars Hernquist, Federico Marinacci, Rohan P. Naidu, Melissa K. Ness, Minjung Park, Vadim A. Semenov, and Turner Woody for helpful discussions. We would like to thank Marc Pinsonneault for sharing early access to the APOKASC-3 dataset. We would like to thank Filippo Barbani for kindly providing his version of \texttt{MakeNewDisk}. A.B. would like to thank Todd Phillips for helpful discussions.
